---
title: "final_project_censusdata"
output: html_notebook
---

```{r}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(tigris)
library(sf)
library(tmap)
library(leaflet)
library(ggiraph)
library(patchwork)
library(shiny)
```

#import crime data
#Link source: https://crime-data-explorer.app.cloud.gov/pages/explorer/crime/crime-trend
```{r}
all_property <- read_csv("data/all_property_crime.csv")
all_violence <- read_csv("data/all_violent.csv")
```

#Variables
```{r}
vars_2020 <- load_variables(2020, "acs5", cache = FALSE)
vars_2015 <- load_variables(2015, "acs5", cache = FALSE)
vars_2010 <- load_variables(2010, "acs5", cache = FALSE)
```

```{r}
profile_2020 <- load_variables(2020, "acs5/profile")
profile_2019 <- load_variables(2019, "acs5/profile")
profile_2018 <- load_variables(2018, "acs5/profile")
profile_2016 <- load_variables(2016, "acs5/profile")
profile_2015 <- load_variables(2015, "acs5/profile")
profile_2010 <- load_variables(2010, "acs5/profile")
```

#compare profile
```{r}
compare_profile_2020 <- load_variables(2020, "acs5/cprofile")
compare_profile_2015 <- load_variables(2015, "acs5/cprofile")
```

#Median age
```{r}
years <- c(2010:2020)

median_age <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(median_age = "B01002_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year) 
})

#How to drop Lake St. Louis city?
median_age %>% filter(grepl("St. Louis", NAME) | grepl("Columbia", NAME) | grepl("Kansas", NAME) | grepl("Lee's", NAME) | grepl("Independence", NAME) | grepl("Springfield", NAME) | grepl("O'Fallon", NAME))

#Columbia median age 2010:2020
columbia_median_age <- median_age %>% filter(grepl("Columbia", NAME))

write_csv(columbia_median_age, "data/columbia_median_age.csv")

```

#Columbia population pyramid

```{r}
subject_2010 <- load_variables(2010, "acs5/subject")
```


```{r}
cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C02_0", .) %>%
  set_names(cohort_names)

```

```{r}
#2010
male_data_2010 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2010
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2010 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2010
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2010 <- bind_rows(male_data_2010, female_data_2010)%>%
  mutate(variable = factor(variable, levels = cohort_names))

write_csv(pyramid_2010, "data/pyramid_2010.csv")
```

```{r}
ggplot(pyramid_2010, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2010 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2006-2010 ACS & tidycensus R package")

ggsave(filename = "pyramid2010.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)
```

```{r}
#2015

subject_2015 <- load_variables(2015, "acs5/subject")

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C02_0", .) %>%
  set_names(cohort_names)

male_data_2015 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2015
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2015 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2015
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2015 <- bind_rows(male_data_2015, female_data_2015)%>%
  mutate(variable = factor(variable, levels = cohort_names))

write_csv(pyramid_2015, "data/pyramid_2015.csv")

```

```{r}
pop_pyramid_2015 <- ggplot(pyramid_2015, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2015 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2011-2015 ACS & tidycensus R package")


ggsave(filename = "pop_pyramid_2015.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)

```


```{r}
#2020

subject_2020 <- load_variables(2020, "acs5/subject")

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C05_0", .) %>%
  set_names(cohort_names)

male_data_2020 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2020
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2020 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2020
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2020 <- bind_rows(male_data_2020, female_data_2020)%>%
  mutate(variable = factor(variable, levels = cohort_names))

write_csv(pyramid_2020, "data/pyramid_2020.csv")

```

```{r}
ggplot(pyramid_2020, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2020 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2016-2020 ACS & tidycensus R package")


ggsave(filename = "columbia_pyramid_2020.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)
```


#Median income
```{r}
years <- c(2010:2020)

median_income <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(median_income = "B19013_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year)
})
  
columbia_median_income <- median_income %>% filter(grepl("Columbia", NAME))

write_csv(columbia_median_income, "data/columbia_median_income.csv")
```


#Poverty rate
```{r}
#Missouri's cities poverty rate in 2020: to see where Columbia is

#B17010_002: Estimate!!Total:!!Income in the past 12 months below poverty level (Concept: POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN)

#B17010_001: Estimate!!Total: (Concept: POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN)

mo_poverty <- get_acs(
  geography = "place",
  variables = c(poverty_num = "B17010_002",
                poverty_denom = "B17010_001"),
  state = "MO",
  year = 2020,
  output = "wide"
) %>% filter(grepl("city", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))

#Because there are alot of small cities, I'll filter places where poverty_denomE is bigger than 20,000
mo_bigcity_poverty_2020 <- mo_poverty %>% filter(poverty_denomE > 20000)

#Columbia is 3/8 among big cities in Missouri in term of lowest poverty percent (8.5%)
```

#Let's see how poverty percent of Columbia has changed the last decade
```{r}
years <- c(2010:2020)

mo_decade_poverty <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(poverty_num = "B17010_002",
                poverty_denom = "B17010_001"),
    state = "MO",
    year = year,
    output = "wide"
  ) %>%  mutate(year = year) %>% 
    filter(grepl("city", NAME)) %>% 
    mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))
})

#Because there are alot of small cities, I'll filter places where poverty_denomE is bigger than 20,000
mo_bigcity_poverty <- mo_decade_poverty %>% filter(poverty_denomE > 20000)


#Columbia decade poverty rate
columbia_decade_poverty <- mo_decade_poverty %>% filter(grepl("Columbia", NAME))

write_csv(columbia_decade_poverty, "data/columbia_decade_poverty.csv")
#Good news: The poverty percent has declined year by year, hit the lowest in 2020

```

#Let see Columbia's poverty ranking year to year
```{r}
#How to do a loop here?

mo_bigcity_poverty %>% select(NAME, year, percent) %>% 
  pivot_wider(names_from = year, values_from = percent)

#2010
mo_bigcity_poverty %>% filter(year == 2010) %>% 
  arrange(percent)
#Ranked 2nd lowest poverty percent among 8 big cities with 11%

#2011
mo_bigcity_poverty %>% filter(year == 2011) %>% 
  arrange(percent)

#Ranked 3rd, 10.4%

#2012
mo_bigcity_poverty %>% filter(year == 2012) %>% 
  arrange(percent)
```


```{r}
#I think it makes sense more if compare Columbia to Missouri's Cosmopolitan Area
mo_cbsa_poverty <- get_acs(
    geography = "cbsa",
    variables = c(poverty_num = "B17010_002",
                  poverty_denom = "B17010_001"),
    year = 2020, 
    output = "wide"
  ) %>% 
  filter(grepl("MO", NAME) & grepl("Metro Area", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))
  
#Columbia ranked 5/8 of Missouri's Metro Area with 7.68% of poverty
```

```{r}
years <- c(2010:2020)

cbsa_10year_poverty <- map_dfr(years, function(year) {
  get_acs(
    geography = "cbsa",
    variables = c(poverty_num = "B17010_002",
                  poverty_denom = "B17010_001"),
    year = year,
    output = "wide"
  ) %>% 
  mutate(year = year)
}) 

cbsa_10year_poverty %>% filter(grepl("MO", NAME) & grepl("Metro Area", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  select(NAME, year, percent) %>% 
  pivot_wider(names_from = year, values_from = percent)
```
# Education
```{r}
#CP02PR_2015_068: 2011-2015 Estimates!!EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher (from compare_profile_2020)
#No data available?
get_acs(
  geography = "county",
  table = "CP02PR_2015",
  state = "MO",
  year = 2020
)
```
#Note for education: Inconsistent variables (refer to profile series)
2010-2018: DP02_0067: Estimate!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher
            DP02_0067P: Percent!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher

2019-2020: DP02_0068: Estimate!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher
           DP02_0068P:Percent!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher

```{r}
years <- c(2010:2018)

education1 <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(percent = "DP02_0067P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_education1 <- education1 %>% filter(grepl("Columbia", NAME))

year1 <- c(2019:2020)

education2 <- map_dfr(year1, function(year) {
  get_acs(
    geography = "place",
    variables = c(percent = "DP02_0068P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_education2 <- education2 %>% filter(grepl("Columbia", NAME))

columbia_education <- bind_rows(columbia_education1, columbia_education2)

write_csv(columbia_education, "data/columbia_education.csv")
```

#Unemployment
```{r}
#DP03_0009P: Percent!!EMPLOYMENT STATUS!!Civilian labor force!!Unemployment Rate

years <- c(2010:2020)

unemployment <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(unemployment_rate = "DP03_0009P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_unemployment <- unemployment %>% filter(grepl("Columbia", NAME))

write_csv(columbia_unemployment, "data/columbia_unemployment.csv")
```

#Columbia map in 2011 and 2020 (places function is not currently available for years prior to 2011)

```{r}
mo_places <- places("MO", cb = FALSE, year = 2020)

columbia <- mo_places %>% filter(grepl("Columbia", NAME))

ggplot(columbia) +
  geom_sf() +
  theme_void() +
  labs(title = "Columbia 2020")
```

```{r}
mo_places_2011 <- places("MO", cb = FALSE, year = 2011)

columbia2011 <- mo_places_2011 %>% filter(grepl("Columbia", NAME))

ggplot(columbia2011) +
  geom_sf() +
  theme_void() +
  labs(title = "Columbia 2011")
```
#MO median_age by counties
```{r}
mo_counties_median_age <- get_acs(
  geography = "county",
  variables = "B01002_001",
  state = "MO",
  year = 2020,
  geometry = TRUE,
  resolution = "20m")

mo_counties_median_age %>% ggplot(
  aes(fill = estimate)) +
  geom_sf() +
  scale_fill_distiller(palette = "PuBu",
                       direction = 1) +
  labs(title = "Missouri Median Age by County, 2020",
       caption = "Data source: 2020 5-year ACS, US Census Bureau",
       fill = "ACS estimate") +
  theme_void()

ggsave(filename = "missouri_median_age.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)

```
#MO top 20 counties with highest income
```{r}
mo_income_county <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "MO",
  year = 2020,
  geometry = TRUE
) %>% mutate(NAME = str_remove(NAME, " County, Missouri"))

mo_income_county <- mo_income_county %>% arrange(desc(estimate))

mo_top20 <- mo_income_county %>% filter(estimate > 58426)
```

```{r}
mo_map <- ggplot(mo_top20, aes(fill = estimate)) +
  geom_sf_interactive(aes(data_id = GEOID)) +
  scale_fill_distiller(palette = "Greens",
                       direction = 1,
                       guide = "none") +
  theme_void()

mo_plot <- ggplot(mo_top20, aes(x = estimate, y = reorder(NAME, estimate), fill = estimate)) +
  geom_errorbarh(aes(xmin = estimate - moe, xmax = estimate + moe)) +
  geom_point_interactive(color = "black", size = 4, shape = 21,
                         aes(data_id = GEOID)) +
  scale_fill_distiller(palette = "Greens", direction = 1,
                       labels = scales::dollar) +
  scale_x_continuous(labels = scales::dollar) +
  labs(title = "Top 20 household income by county in Missouri",
       subtitle = "2016 - 2020 American Community Survey",
       y = "",
       x = "ACS estimate (bars represent margin of error)",
       fill = "ACS estimate") +
  theme_minimal(base_size = 14)

girafe(ggobj = mo_map + mo_plot, width_svg = 10, height_svg = 5) %>% 
  girafe_options(opts_hover(css = "fill:cyan;"))
```

#top 20 highest income counties in 2010
```{r}
mo_income_county2010 <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "MO",
  year = 2010,
  geometry = TRUE
) %>% mutate(NAME = str_remove(NAME, " County, Missouri"))

mo_income_county2010 <- mo_income_county2010 %>% arrange(desc(estimate))

mo_top20_2010 <- mo_income_county2010 %>% filter(estimate > 45746)
```

```{r}
mo_map_2010 <- ggplot(mo_top20_2010, aes(fill = estimate)) +
  geom_sf_interactive(aes(data_id = GEOID)) +
  scale_fill_distiller(palette = "Greens",
                       direction = 1,
                       guide = "none") +
  theme_void()

mo_plot_2010 <- ggplot(mo_top20_2010, aes(x = estimate, y = reorder(NAME, estimate), fill = estimate)) +
  geom_errorbarh(aes(xmin = estimate - moe, xmax = estimate + moe)) +
  geom_point_interactive(color = "black", size = 4, shape = 21,
                         aes(data_id = GEOID)) +
  scale_fill_distiller(palette = "Greens", direction = 1,
                       labels = scales::dollar) +
  scale_x_continuous(labels = scales::dollar) +
  labs(title = "Top 20 household income by county in Missouri",
       subtitle = "2006 - 2010 American Community Survey",
       y = "",
       x = "ACS estimate (bars represent margin of error)",
       fill = "ACS estimate") +
  theme_minimal(base_size = 14)

girafe(ggobj = mo_map_2010 + mo_plot_2010, width_svg = 10, height_svg = 5) %>% 
  girafe_options(opts_hover(css = "fill:cyan;"))
```


#Race 
```{r}
twin_cities_race <- get_acs(
  geography = "tract",
  variables = c(
    hispanic = "DP05_0071P",
    white = "DP05_0077P",
    black = "DP05_0078P",
    native = "DP05_0079P",
    asian = "DP05_0080P"),
  state = "MO",
  county = c("Boone", "Audrain", "Randolph", "Cooper",
             "Howard"),
  year = 2020,
  geometry = TRUE) 

groups <- c("Hispanic" = "hispanic",
            "White" = "white",
            "Black" = "black",
            "Native American" = "native",
            "Asian" = "asian")

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput(
        inputId = "group",
        label = "Select a group to map",
        choices = groups
      )
    ),
    mainPanel(
      leafletOutput("map", height = "600")
    )
  )
)

server <- function(input, output) {
  
  # Reactive function that filters for the selected group in the drop-down menu
  group_to_map <- reactive({
    filter(twin_cities_race, variable == input$group)
  })
  
  # Initialize the map object, centered on the Minneapolis-St. Paul area
  output$map <- renderLeaflet({

    leaflet(options = leafletOptions(zoomControl = FALSE)) %>%
      addProviderTiles(providers$Stamen.TonerLite) %>%
      setView(lng = -93.21,
              lat = 44.98,
              zoom = 8.5)

  })
  
  observeEvent(input$group, {
    
    pal <- colorNumeric("viridis", group_to_map()$estimate)
    
    leafletProxy("map") %>%
      clearShapes() %>%
      clearControls() %>%
      addPolygons(data = group_to_map(),
                  color = ~pal(estimate),
                  weight = 0.5,
                  fillOpacity = 0.5,
                  smoothFactor = 0.2,
                  label = ~estimate) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = group_to_map()$estimate,
        title = "% of population"
      )
  })
  
}

shinyApp(ui = ui, server = server)
```


# go back to 1950 - 2000
```{r}

```


