---
title: "final_project_censusdata"
output: html_notebook
---
#Use tidycensus package to pull data from US Census Bureau to see how City of Columbia has changed over the past decade (2011 - 2020) 
# Source: Kyle Walker's tidycensus: https://github.com/walkerke/tidycensus and his book: https://walker-data.com/census-r/. I can still have a lot of things to learn from this guy and the book. 
#Special thanks to Liz Lucas who has introduced me to R and MaryJo Webster who has introduced me to Kyle Walker.


```{r}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(tigris)
library(sf)
library(tmap)
library(leaflet)
library(ggiraph)
library(patchwork)
library(shiny)
```

# Load Variables to refer
```{r}
vars_2020 <- load_variables(2020, "acs5", cache = FALSE) #American Comunity Survey 5 years variable
vars_2015 <- load_variables(2015, "acs5", cache = FALSE)
vars_2010 <- load_variables(2010, "acs5", cache = FALSE)

den_2010 <- load_variables(2010, "pl", cache = FALSE) #decenial variables 2010
```

#Load profile variables
```{r}
profile_2020 <- load_variables(2020, "acs5/profile")
profile_2019 <- load_variables(2019, "acs5/profile")
profile_2018 <- load_variables(2018, "acs5/profile")
profile_2016 <- load_variables(2016, "acs5/profile")
profile_2015 <- load_variables(2015, "acs5/profile")
profile_2010 <- load_variables(2010, "acs5/profile")
```

#compare profile variables (new function of tidycensus package)
```{r}
compare_profile_2020 <- load_variables(2020, "acs5/cprofile")
compare_profile_2015 <- load_variables(2015, "acs5/cprofile")
```

#Total population 2011 - 2020. Geography = "place"
#Variable: B01001_001 Estimate!!Total:

```{r}
years <- c(2011:2020)

total_population <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(total_population = "B01001_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year) 
}) %>% 
  filter(grepl("Columbia", NAME))

write_csv(total_population, "total_population.csv")

#Columbia population ranking among cities in the U.S
#Ranked: 238
pop_rank <- get_acs(
  geography = "place",
  variables = "B01001_001",
  year = 2020
) %>% arrange(desc(estimate))
```

#Median age
#Variable: B01002_001 Estimate!!Median age --!!Total:
```{r}
years <- c(2011:2020)

median_age <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(median_age = "B01002_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year) 
})

#Columbia median age 2011:2020
columbia_median_age <- median_age %>% filter(grepl("Columbia", NAME))

write_csv(columbia_median_age, "data/columbia_median_age.csv")

#Ranked 2203/31908 "places" in the State in 2020
median_age_rank <- get_acs(
  geography = "place",
  variables = "B01002_001",
  year = 2020
) %>% arrange(estimate)

#Ranked 83/1082 place of MO
median_age_rank_mo <- get_acs(
  geography = "place",
  state = "MO",
  variables = "B01002_001",
  year = 2020
) %>% arrange(estimate)

```

#Columbia population pyramid 2010 - 2015 - 2020 to see how the population components have changed. 
#Load subject variables
```{r}
subject_2010 <- load_variables(2010, "acs5/subject")
```


```{r}
#Variables S0101_C03_0xx are female variables with different age groups. 2:19 are the values to paste in xx position, as age groups

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

#Variables S0101_C02_0xx are male variables with different age groups

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C02_0", .) %>%
  set_names(cohort_names)

```

```{r}
#2010
male_data_2010 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2010
) %>% filter(grepl("Columbia", NAME)) %>% #Filter Columbia population
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% #Remove unnecessary words to clean the data
  mutate(sex = "Male",
         estimate = estimate * -1) #-1 to make male chart on the opposite side of female 


female_data_2010 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2010
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2010 <- bind_rows(male_data_2010, female_data_2010)%>%
  mutate(variable = factor(variable, levels = cohort_names))

write_csv(pyramid_2010, "data/pyramid_2010.csv")
```

```{r}
#Visualize it by ggplot2
ggplot(pyramid_2010, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + #I choose Mizzou's colors
  labs(x = "", 
       y = "ACS estimate", 
       title = "2010 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2006-2010 ACS & tidycensus R package") #Credit 

ggsave(filename = "pyramid2010.png", #Save the pyramid
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)
```

```{r}
#2015

subject_2015 <- load_variables(2015, "acs5/subject")

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C02_0", .) %>%
  set_names(cohort_names)

male_data_2015 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2015
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2015 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2015
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2015 <- bind_rows(male_data_2015, female_data_2015)%>% #rbind the data 
  mutate(variable = factor(variable, levels = cohort_names))

write_csv(pyramid_2015, "data/pyramid_2015.csv")

```

```{r}
pop_pyramid_2015 <- ggplot(pyramid_2015, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2015 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2011-2015 ACS & tidycensus R package")


ggsave(filename = "pop_pyramid_2015.png", # Save the pyramid
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)

```


```{r}
#2020

subject_2020 <- load_variables(2020, "acs5/subject")

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C05_0", .) %>%
  set_names(cohort_names)

male_data_2020 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2020
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2020 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2020
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2020 <- bind_rows(male_data_2020, female_data_2020)%>%
  mutate(variable = factor(variable, levels = cohort_names))

write_csv(pyramid_2020, "data/pyramid_2020.csv")

```

```{r}
ggplot(pyramid_2020, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2020 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2016-2020 ACS & tidycensus R package")


ggsave(filename = "columbia_pyramid_2020.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)
```
#Columbia population is getting slightly older, but it hasn't changed much the last decade. As a college town, young students have claimed a large part of the city's population and contribute not not much change population's structure.

#Median income
#B19013_001 Estimate!!Median household income in the past 12 months
```{r}
years <- c(2011:2020)

median_income <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(median_income = "B19013_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year)
})
  
columbia_median_income <- median_income %>% filter(grepl("Columbia", NAME))

write_csv(columbia_median_income, "data/columbia_median_income.csv")

#Ranked 14525/31908 places in the US, but the ranking doesn't mean much, since the estimate value for 
#small geography level like place has large margin of error (especially with the places that have small sample size)
median_income_rank <- get_acs(
  geography = "place",
  variables = "B19013_001",
  year = 2020
) %>% arrange(desc(estimate))

#Ranked 310/1082 places of MO
median_income_rank_mo <- get_acs(
  geography = "place",
  state = "MO",
  variables = "B19013_001",
  year = 2020
) %>% arrange(desc(estimate))
```


#Poverty rate
```{r}
#Missouri's cities poverty rate in 2020: to see where Columbia is

#B17010_002: Estimate!!Total:!!Income in the past 12 months below poverty level (Concept: POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN)

#B17010_001: Estimate!!Total: (Concept: POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN)

mo_poverty <- get_acs(
  geography = "place",
  variables = c(poverty_num = "B17010_002",
                poverty_denom = "B17010_001"),
  state = "MO",
  year = 2020,
  output = "wide"
) %>% filter(grepl("city", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))

#Because there are alot of small cities, I'll filter places where poverty_denomE is bigger than 20,000
mo_bigcity_poverty_2020 <- mo_poverty %>% filter(poverty_denomE > 20000)

#Columbia is 3/8 among big cities in Missouri in term of lowest poverty percent (8.5%)
```

#Let's see how poverty percent of Columbia has changed the last decade
```{r}
years <- c(2011:2020)

mo_decade_poverty <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(poverty_num = "B17010_002",
                poverty_denom = "B17010_001"),
    state = "MO",
    year = year,
    output = "wide"
  ) %>%  mutate(year = year) %>% 
    filter(grepl("city", NAME)) %>% 
    mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))
})

#Because there are alot of small cities, I'll filter places where poverty_denomE is bigger than 20,000
mo_bigcity_poverty <- mo_decade_poverty %>% filter(poverty_denomE > 20000)


#Columbia decade poverty rate
columbia_decade_poverty <- mo_decade_poverty %>% filter(grepl("Columbia", NAME))

ggplot(columbia_decade_poverty, aes(x = year, y = percent)) + 
  geom_line(color = "navy") + 
  geom_point(color = "navy", size = 2) + 
  theme_minimal(base_size = 12) + 
  labs(title = "Columbia poverty percent over last decade",
       x = "Year",
       y = "percent") +
  scale_y_continuous(limits = c(0, 12))

write_csv(columbia_decade_poverty, "data/columbia_decade_poverty.csv")
#Good news: The poverty percent has declined year by year, hit the lowest in 2020

```



```{r}
#I think it makes sense more if compare Columbia to Missouri's Cosmopolitan Area
mo_cbsa_poverty <- get_acs(
    geography = "cbsa",
    variables = c(poverty_num = "B17010_002",
                  poverty_denom = "B17010_001"),
    year = 2020, 
    output = "wide"
  ) %>% 
  filter(grepl("MO", NAME) & grepl("Metro Area", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))
  
#In 2020, Columbia ranked 5/8 of Missouri's Metro Area with 7.68% of poverty
```

```{r}
#If we want further exploration of poverty rate among all metro area in the U.S, we can pull the data for the last decade and examine it. In this script, I'm more on comparing Columbia with itself in the past than compare with other places. Just pull the data for later useage.
years <- c(2011:2020)

cbsa_10year_poverty <- map_dfr(years, function(year) {
  get_acs(
    geography = "cbsa",
    variables = c(poverty_num = "B17010_002",
                  poverty_denom = "B17010_001"),
    year = year,
    output = "wide"
  ) %>% 
  mutate(year = year)
}) 

cbsa_10year_poverty %>% filter(grepl("MO", NAME) & grepl("Metro Area", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  select(NAME, year, percent) %>% 
  pivot_wider(names_from = year, values_from = percent)
```
# Education

#Note: Inconsistent variables (refer to profile series)
2011-2018: DP02_0067: Estimate!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher
            DP02_0067P: Percent!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher

2019-2020: DP02_0068: Estimate!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher
           DP02_0068P:Percent!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher

```{r}
years <- c(2011:2018)

education1 <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(percent = "DP02_0067P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_education1 <- education1 %>% filter(grepl("Columbia", NAME))

year1 <- c(2019:2020)

education2 <- map_dfr(year1, function(year) {
  get_acs(
    geography = "place",
    variables = c(percent = "DP02_0068P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_education2 <- education2 %>% filter(grepl("Columbia", NAME))

columbia_education <- bind_rows(columbia_education1, columbia_education2)

write_csv(columbia_education, "data/columbia_education.csv")

#Ranked 180/1082?
edu_rank <- get_acs(
    geography = "place",
    state = "MO",
    variables = "DP02_0067P",
    year = 2020) %>% arrange(desc(estimate))
```

#Unemployment
```{r}
#DP03_0009P: Percent!!EMPLOYMENT STATUS!!Civilian labor force!!Unemployment Rate

years <- c(2010:2020)

unemployment <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(unemployment_rate = "DP03_0009P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_unemployment <- unemployment %>% filter(grepl("Columbia", NAME))

write_csv(columbia_unemployment, "data/columbia_unemployment.csv")
```

#This data is to examine amount of people not in labor force
```{r}
#DP03_0007P Percent!!EMPLOYMENT STATUS!!Population 16 years and over!!Not in labor force
years <- c(2011:2020)

notin_laborforce <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(unemployment_rate = "DP03_0007P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year) 
}) %>% filter(grepl("Columbia", NAME))

write_csv(notin_laborforce, "notin_laborforce.csv")
```

#Columbia map in 2011 and 2020 (places function is not currently available for years prior to 2011). Just to see how the city has changed geographically.

```{r}
mo_places <- places("MO", cb = FALSE, year = 2020)

columbia <- mo_places %>% filter(grepl("Columbia", NAME))

ggplot(columbia) +
  geom_sf() +
  theme_void() +
  labs(title = "Columbia 2020")
```

```{r}
mo_places_2011 <- places("MO", cb = FALSE, year = 2011)

columbia2011 <- mo_places_2011 %>% filter(grepl("Columbia", NAME))

ggplot(columbia2011) +
  geom_sf() +
  theme_void() +
  labs(title = "Columbia 2011")
```

