---
title: "final_project_censusdata"
output: html_notebook
---

```{r}
library(tidycensus)
library(tidyverse)
library(ggplot2)
```

#Variables
```{r}
vars_2020 <- load_variables(2020, "acs5", cache = FALSE)
vars_2015 <- load_variables(2015, "acs5", cache = FALSE)
vars_2010 <- load_variables(2010, "acs5", cache = FALSE)
```

```{r}
profile_2020 <- load_variables(2020, "acs5/profile")
profile_2019 <- load_variables(2019, "acs5/profile")
profile_2018 <- load_variables(2018, "acs5/profile")
profile_2016 <- load_variables(2016, "acs5/profile")
profile_2015 <- load_variables(2015, "acs5/profile")
profile_2010 <- load_variables(2010, "acs5/profile")
```

#compare profile
```{r}
compare_profile_2020 <- load_variables(2020, "acs5/cprofile")
compare_profile_2015 <- load_variables(2015, "acs5/cprofile")
```

#Median age
```{r}
years <- c(2010:2020)

median_age <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(median_age = "B01002_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year) 
})

#How to drop Lake St. Louis city?
median_age %>% filter(grepl("St. Louis", NAME) | grepl("Columbia", NAME) | grepl("Kansas", NAME) | grepl("Lee's", NAME) | grepl("Independence", NAME) | grepl("Springfield", NAME) | grepl("O'Fallon", NAME))

#Columbia median age 2010:2020
columbia_median_age <- median_age %>% filter(grepl("Columbia", NAME))

write_csv(columbia_median_age, "data/columbia_median_age.csv")
```

#Columbia population pyramid

```{r}
subject_2010 <- load_variables(2010, "acs5/subject")
```


```{r}
cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C02_0", .) %>%
  set_names(cohort_names)

```

```{r}
#2010
male_data_2010 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2010
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2010 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2010
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2010 <- bind_rows(male_data_2010, female_data_2010)%>%
  mutate(variable = factor(variable, levels = cohort_names))
```

```{r}
ggplot(pyramid_2010, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2010 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2006-2010 ACS & tidycensus R package")

ggsave(filename = "pyramid2010.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)
```

```{r}
#2015

subject_2015 <- load_variables(2015, "acs5/subject")

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C02_0", .) %>%
  set_names(cohort_names)

male_data_2015 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2015
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2015 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2015
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2015 <- bind_rows(male_data_2015, female_data_2015)%>%
  mutate(variable = factor(variable, levels = cohort_names))

```

```{r}
ggplot(pyramid_2015, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2015 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2011-2015 ACS & tidycensus R package")


ggsave(filename = "columbia_pyramid_2015.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)

```


```{r}
#2020

subject_2020 <- load_variables(2020, "acs5/subject")

cohort_names <- c("0-4", "5-9", "10-14", "15-19",
                  "20-24", "25-29", "30-34", "35-39",
                  "40-44", "45-49", "50-54", "55-59",
                  "60-64", "65-69", "70-74", "75-79",
                  "80-84", "85+")

male_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C03_0", .) %>%
  set_names(cohort_names)

female_vars <- 2:19 %>%
  str_pad(2, "left", "0") %>%
  paste0("S0101_C05_0", .) %>%
  set_names(cohort_names)

male_data_2020 <- get_acs(
  geography = "place",
  state = "MO",
  variables = male_vars,
  year = 2020
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Male",
         estimate = estimate * -1)


female_data_2020 <- get_acs(
  geography = "place",
  state = "MO",
  variables = female_vars,
  year = 2020
) %>% filter(grepl("Columbia", NAME)) %>% 
  mutate(NAME = str_remove(NAME, " city, Missouri")) %>% 
  mutate(sex = "Female")

pyramid_2020 <- bind_rows(male_data_2020, female_data_2020)%>%
  mutate(variable = factor(variable, levels = cohort_names))

```

```{r}
ggplot(pyramid_2020, 
       aes(x = estimate, 
           y = variable, 
           fill = sex)) + 
  geom_col(width = 0.95, alpha = 0.75) + 
  theme_minimal(base_size = 16) + 
  scale_x_continuous(labels = function(x) paste0(abs(x / 1000), "k")) + 
  scale_fill_manual(values = c("#000000", "#F1B82D")) + 
  labs(x = "", 
       y = "ACS estimate", 
       title = "2020 Population structure in Columbia, Missouri", 
       fill = "", 
       caption = "Data source: 2016-2020 ACS & tidycensus R package")


ggsave(filename = "columbia_pyramid_2020.png",
       path = "maps",
       width = 8,
       height = 5,
       units = "in",
       dpi = 300)
```


#Median income
```{r}
years <- c(2010:2020)

median_income <- map_dfr(years, function(year) {
  get_acs(
  geography = "place",
  variables = c(median_income = "B19013_001"),
  state = "MO",
  year = year) %>% 
    mutate(year = year)
})
  
columbia_median_income <- median_income %>% filter(grepl("Columbia", NAME))

write_csv(columbia_median_income, "data/columbia_median_income.csv")
```


#Poverty rate
```{r}
#Missouri's cities poverty rate in 2020: to see where Columbia is

#B17010_002: Estimate!!Total:!!Income in the past 12 months below poverty level (Concept: POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN)

#B17010_001: Estimate!!Total: (Concept: POVERTY STATUS IN THE PAST 12 MONTHS OF FAMILIES BY FAMILY TYPE BY PRESENCE OF RELATED CHILDREN UNDER 18 YEARS BY AGE OF RELATED CHILDREN)

mo_poverty <- get_acs(
  geography = "place",
  variables = c(poverty_num = "B17010_002",
                poverty_denom = "B17010_001"),
  state = "MO",
  year = 2020,
  output = "wide"
) %>% filter(grepl("city", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))

#Because there are alot of small cities, I'll filter places where poverty_denomE is bigger than 20,000
mo_bigcity_poverty_2020 <- mo_poverty %>% filter(poverty_denomE > 20000)

#Columbia is 3/8 among big cities in Missouri in term of lowest poverty percent (8.5%)
```

#Let's see how poverty percent of Columbia has changed the last decade
```{r}
years <- c(2010:2020)

mo_decade_poverty <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(poverty_num = "B17010_002",
                poverty_denom = "B17010_001"),
    state = "MO",
    year = year,
    output = "wide"
  ) %>%  mutate(year = year) %>% 
    filter(grepl("city", NAME)) %>% 
    mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))
})

#Because there are alot of small cities, I'll filter places where poverty_denomE is bigger than 20,000
mo_bigcity_poverty <- mo_decade_poverty %>% filter(poverty_denomE > 20000)


#Columbia decade poverty rate
columbia_decade_poverty <- mo_decade_poverty %>% filter(grepl("Columbia", NAME))

write_csv(columbia_decade_poverty, "data/columbia_decade_poverty.csv")
#Good news: The poverty percent has declined year by year, hit the lowest in 2020

```

#Let see Columbia's poverty ranking year to year
```{r}
#How to do a loop here?

mo_bigcity_poverty %>% select(NAME, year, percent) %>% 
  pivot_wider(names_from = year, values_from = percent)

#2010
mo_bigcity_poverty %>% filter(year == 2010) %>% 
  arrange(percent)
#Ranked 2nd lowest poverty percent among 8 big cities with 11%

#2011
mo_bigcity_poverty %>% filter(year == 2011) %>% 
  arrange(percent)

#Ranked 3rd, 10.4%

#2012
mo_bigcity_poverty %>% filter(year == 2012) %>% 
  arrange(percent)
```


```{r}
#I think it makes sense more if compare Columbia to Missouri's Cosmopolitan Area
mo_cbsa_poverty <- get_acs(
    geography = "cbsa",
    variables = c(poverty_num = "B17010_002",
                  poverty_denom = "B17010_001"),
    year = 2020, 
    output = "wide"
  ) %>% 
  filter(grepl("MO", NAME) & grepl("Metro Area", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  arrange(desc(percent))
  
#Columbia ranked 5/8 of Missouri's Metro Area with 7.68% of poverty
```

```{r}
years <- c(2010:2020)

cbsa_10year_poverty <- map_dfr(years, function(year) {
  get_acs(
    geography = "cbsa",
    variables = c(poverty_num = "B17010_002",
                  poverty_denom = "B17010_001"),
    year = year,
    output = "wide"
  ) %>% 
  mutate(year = year)
}) 

cbsa_10year_poverty %>% filter(grepl("MO", NAME) & grepl("Metro Area", NAME)) %>% 
  mutate(percent = (poverty_numE/poverty_denomE)*100) %>% 
  select(NAME, year, percent) %>% 
  pivot_wider(names_from = year, values_from = percent)
```
# Education
```{r}
#CP02PR_2015_068: 2011-2015 Estimates!!EDUCATIONAL ATTAINMENT!!Population 25 years and over!!Bachelor's degree or higher (from compare_profile_2020)
#No data available?
get_acs(
  geography = "county",
  table = "CP02PR_2015",
  state = "MO",
  year = 2020
)
```
#Note for education: Inconsistent variables (refer to profile series)
2010-2018: DP02_0067: Estimate!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher
            DP02_0067P: Percent!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher

2019-2020: DP02_0068: Estimate!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher
           DP02_0068P:Percent!!EDUCATIONAL ATTAINMENT!!Percent bachelor's degree or higher

```{r}
years <- c(2010:2018)

education1 <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(percent = "DP02_0067P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_education1 <- education1 %>% filter(grepl("Columbia", NAME))

year1 <- c(2019:2020)

education2 <- map_dfr(year1, function(year) {
  get_acs(
    geography = "place",
    variables = c(percent = "DP02_0068P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_education2 <- education2 %>% filter(grepl("Columbia", NAME))

columbia_education <- bind_rows(columbia_education1, columbia_education2)

write_csv(columbia_education, "data/columbia_education.csv")
```

#Unemployment
```{r}
#DP03_0009P: Percent!!EMPLOYMENT STATUS!!Civilian labor force!!Unemployment Rate

years <- c(2010:2020)

unemployment <- map_dfr(years, function(year) {
  get_acs(
    geography = "place",
    variables = c(unemployment_rate = "DP03_0009P"),
    state = "MO",
    year = year
  ) %>% mutate(year = year)
})

columbia_unemployment <- unemployment %>% filter(grepl("Columbia", NAME))

write_csv(columbia_unemployment, "data/columbia_unemployment.csv")
```

